---
title: "Arrival Delay"
author: "Mohan Kandaraj"
date: "October 4, 2015"
output: html_document
---
R program to tidy up a wide data for analysis

####Wide Data
```{r, message=FALSE}
delay_wide=read.csv(file="https://raw.githubusercontent.com/mkds/MSDA/master/IS607/data/delays.csv", na.strings = "")
delay_wide
```

####Tidy up Data   
Let's transform the data that will allow as to compare the delays of two airlines  
 
```{r, message=FALSE, warning=FALSE}
library(tidyr)
library(dplyr)
library(reshape2)


delay <- delay_wide  %>%
  gather(airport,flights,3:7,na.rm=T) %>% #Convert Airport from multi column to a column value pair
  fill(1) %>% #If a value is blank then fill with non blank value from previous row
  spread(X.1,flights) %>% #Convert one column key value to multi column value
  rename(airline=X) %>% #Rename column X to airline
  melt %>% 
  dcast(airline + airport ~ variable,sum,margins="airport") %>% #Melt and Recast to add totals by airline
  mutate(airport=gsub("\\."," ",airport),delay_pct=delayed/(delayed+`on time`))
   
#Let's look at transformed Data
delay
```

####Analysis of Delay 

From below chart we could see that at every airport the delay percentage for "AM WEST"" is higher than ALASKA airline. But if we look at all flights (the first bar on the chart) then ALASKA has more arrival delays than AM WEST. So who is the winner?

```{r plot, message=FALSE, warning=FALSE}
library(ggplot2)
ggplot(delay,aes(x=airport,y=delay_pct))+geom_bar(aes(fill=airline),stat="identity",position="dodge")+geom_text(aes(label=format(100*..y.., digits=2)),stat="identity",vjust=-.5,hjust=.5)+labs(x="Airport",y="Delay",fill="Airline") 
```
