---
title: "Income Analysis"
author: "Mohan Kandaraj"
date: "September 12, 2015"
output: 
  html_document:
          toc: true
---
---
###About 
This page provides simple analysis of census data. The dataset used in the page is sourced from <https://archive.ics.uci.edu/ml/datasets/Adult>. 

###Load Data
Let's read and examine the data.
```{r Load}
file_url <- "https://archive.ics.uci.edu/ml/machine-learning-databases/adult/adult.data"
income=read.table(file=file_url,sep=",")
str(income)
```
Name the variables
```{r name}
colnames(income)<-c("age","workclass","fnlwgt","education","education_num","marital_status","occupation","relationship","race","sex","capital_gain","capital_loss","hours_per_week","native_country","income_level")
```

###Tidy Up and Prepare Data
Let's tidy up the data

```{r}
#Drop columns that are not required
income$fnlwgt=NULL
#Attach the dataframe for easy reference
attach(income)
#Check for missing values that may cause problem. Dataset has unknown values i.e a level for ?. But that is fine as it would just be considered as a category
table(complete.cases(income))
#There are no missing values
#Is capital-gain and capital-loss are exclusive? If not net gain/loss would be useful
any(capital_gain!=0 & capital_loss!=0)
#Let's create a variable which captures if the person is involved in capital market using the capital gain and loss variable
income$capital_market = capital_gain !=0 | capital_loss != 0

```

### Visualization

Let's plot some graph to better understand the data and draw some information

Let's check education level

```{r, education}
boxplot(education_num~income_level,outline=F,xlab="Income Level",ylab="Education Level",main="Income Vs Education")

```

Let's check if participation in capital market
```{r, Capital}
prop.table(table(income_level,income$capital_market),2)
# 57.9% people participating in capital market earn more than 50k
```

Let's Plot all categorical variables... 
```{r, Features}
library(ggplot2)
for  (i in 1:ncol(income)-2) {
  if (is.factor(income[,i])){
    pl =ggplot(income,aes_string(colnames(income)[i],fill="income_level"))+geom_bar(position="dodge") + theme(axis.text.x=element_text(angle=70))
    print(pl)
  }
  
}
```

