---
title: "Parse Chess Tournament Data"
author: "Mohan Kandaraj"
output:
  html_document:
    theme: cerulean
    toc: yes
    toc_depth: 4
  word_document: default
---
###About
This page demonstrates text file processing in R using regular expressions.

###Problem Statement
A text file with chess tournament results is given. Generates a .CSV file with the following information for all of the players:
Player's Name, Player's State, Total Number of Points, Player's Pre-Rating, and Average Pre Chess Rating of Opponents

###Solution Approach
Extract the required details from the text file using regular expression. Process the pair of rows that provides the details for a player and store the extracted detail into a row of dataframe. Then, construct the details required for the output CSV file using dataframe operations. Write the output file.

###Solution

Load required libraries 

```{r library, results="hide"}
library(RCurl)
library(stringr)
library(knitr)
```
####Function to parse pair of line
```{r parse_lines}
parse_lines=function(line1_string,line2_string){
  #Reg exp for (id)|(name)|(pts)|(game_string)
  line_one_pattern="(\\d+).+?\\|(.+?)\\|(\\d*\\.?\\d*).+?\\|(.+)"
  #Get required text using Reg exp
  parse1=str_match_all(line1_string,line_one_pattern)[[1]]
 
  #Convert and Store text in dataframe
  value=data.frame(id=as.integer(parse1[2]),name=str_trim(parse1[3]),pts=as.numeric(parse1[4]))
  game_string=parse1[5]
  #Pattern of game details  
  game_pattern="\\D +(\\d*)\\|"
  #Get the opponent list
  opponent_list=as.numeric(str_match_all(game_string,game_pattern)[[1]][,2])
  #Add each opponent as column
  value=cbind(value,matrix(unlist(opponent_list),nrow=1))
  #Name opponent columns properly
  colnames(value)[4:10]=paste0("opponent.",1:7)
  
  #Let's process second line
  #Second line pattern
  line_two_pattern=" +(\\D+).+?\\|.+?: +(\\d+).+"
  #Get required value using reg Exp
  parse2=str_match_all(line2_string,line_two_pattern)[[1]]
  #Add state and pre tournament point of the player
  value=cbind(value,state=parse2[2],pre_point=as.numeric(parse2[3]))
  #Done
  return(value)
}

```
   
Vectorize the function
```{r vector_function}
parse_lines_all=Vectorize(parse_lines,SIMPLIFY = F)
```

####Function that computes opponent score
```{r avg_oppo_rating}
avg_opp_rating=function(x_opp,all_x) round(mean(all_x$pre_point[all_x$id %in% x_opp],na.rm=T))
```
####Main Process
Read the file and call our parse function
```{r}
#Get file data from Github
file_content=getURL("https://raw.githubusercontent.com/mkds/MSDA/master/IS607/data/tournamentinfo.txt")
#Read file data into a dataframe
file_text=read.table(text=file_content,sep="\n",stringsAsFactors = F,skip=4)
#Process in block of three
block_start=seq(1,nrow(file_text),3)
#Pass pair of rows to our parse function and append return values
parsed_value=do.call("rbind",parse_lines_all(file_text[block_start,],file_text[block_start+1,]))
rownames(parsed_value)=NULL

#Compute opponent score
parsed_value$avg_opp_rating=apply(parsed_value[,4:10],1,avg_opp_rating,all_x=parsed_value)
#Select the columns 
output_file=parsed_value[c("name","state","pts","pre_point","avg_opp_rating")]
# Ta da...
kable(output_file)
```
   
   
---
###Output File
Write to File. File Link <https://raw.githubusercontent.com/mkds/MSDA/master/IS607/data/chess_output.csv>
```{r,eval=FALSE}
write.table(output_file,"chess_output.csv",col.names=F,row.names = F,sep=",")
```


