---
title: "Tidy Data - Three Approaches"
author: "Mohan Kandaraj"
output:
  html_document:
    theme: cerulean
    toc: yes
    toc_depth: 4
  word_document: default
---
##Convert untidy Data to Tidy Data - Three approaches   

##Intro   

This page demonstrates three different ways of tidying up the data.   
1. Natural way - Using Tidyr package  
2. Fast(in terms of runtime) way - Using data.table  
3. Hard way - Without using any built-in function/package  

Most of the operations are explained as code comments within the code chunck.

###Natural Way - Using Tidyr Package
I choose Jason Joseph's problem as that was a good wide data candidate for Tidyr package.

Jason has extracted the data from http://www.census.gov/foreign-trade/data/index.html and provided a nice sample file. Thanks Jason for the csv.

```{r library,message=F,warning=F}
library(tidyr)
library(dplyr)
library(knitr)
trade_wide=read.table(file="https://raw.githubusercontent.com/mkds/MSDA/master/IS607/data/International Trade in Goods and Services.csv",stringsAsFactors = F,header = T,na.strings = "",sep = ",")
print(trade_wide)

#Remove Rows which has all the values as NA
trade_wide=trade_wide[rowSums(is.na(trade_wide))!=8,]

#Rename the columns in line with the values of the column
colnames(trade_wide)=c("Month","Type","Exports 2013","Imports 2013","Exports 2014","Imports 2014","Exports 2015","Imports 2015")

#Now let's use tidyr and dplyr to transform the data

trade=trade_wide %>% 
  gather(cat,value,3:8) %>% #Collapse column 3 to 8 into two column (key and value pair)
  separate(cat,c("Trade","year")) %>% #Split the Trade type and year which is stored in one column to two separate columns
  fill(1) %>% #Fill missing values with previous row values
  filter(!is.na(Type)) %>% #Remove rows with type NA
  mutate(value=as.numeric(gsub(",","",value))) #Remove thousand separator in numbers  

head(trade)

#Now that we got the data in desired format, let's answer the question average trade by month and type
kable(trade %>% group_by(Month,Type) %>% summarise(Avg=mean(value)))
```

###Fast way - Using data.table  
For this approach I choose to work with the data provided by Christina Taylor. This data require some transformation and join and so I wanted to try that out in data.table. In general data.table operations tend to execute much faster than tidyr.

```{r,message=F,warning=F}
library(data.table)
library(ggplot2)

jobopening=fread("https://raw.githubusercontent.com/mkds/MSDA/master/IS607/data/JobOpen.csv") #Note: fread is faster than read.csv and does a better job in detecting separator and column classes
head(jobopening)

jobopening[,`Actual Start Date`:=as.Date(`Actual Start Date`,"%m/%d/%Y")] #Convert date string to Date type

#Compute Quater and year of actual start Date
jobopening[,`:=`(Qtr=paste0("Q",(as.numeric(format(`Actual Start Date`,"%m"))-1)%/%3+1),Year=format(`Actual Start Date`,"%Y"))] 

head(jobopening)

hctarget=fread("https://raw.githubusercontent.com/mkds/MSDA/master/IS607/data/hctarget.csv")[-.N]
head(hctarget)

#Collapse Quaterly columns to Qtr/Value column
hctarget=melt(hctarget,measure.vars = c("Q1_2015","Q2_2015","Q3_2015","Q4_2015"),variable.name="Qtr",value.name="Target")

#Split Qtr column to Qtr/Year
hctarget[,c("Qtr","Year"):=tstrsplit(Qtr,"_")]
head(hctarget)

#Set Key values so that we can do data table joins
setkey(jobopening,Role,Location,Qtr,Year)
setkey(hctarget,Role,Location,Qtr,Year)

#Compute total hires and extract only required column
hired=jobopening[!is.na(Year),.(hires=.N),by=.(Role,Location,Qtr,Year)]
#Join hired data and target data
results=hired[hctarget]
#Convert NAs to zero
results$hires[is.na(results$hires)]=0

head(results)
#Heat Map of (Actual Vs Target) by year and Quarter using ggplot facet
ggplot(results,aes(Location, Role)) + geom_tile(aes(fill=hires/(Target-Current_HeadCount)*100)) + scale_fill_gradient2(low="white",high="green") + facet_grid(~Year+Qtr) + labs(fill="Achieved%") + theme(axis.text.x=element_text(size=12),axis.text.y=element_text(size=12))

```

###Hard way - Without using any built-in function/package
What if the data that we get is so messy and we couldn't use any package? I wanted to try and tidy up data without using any package and see how it goes. I choose the data posted by Maxwell Wagner Th data is sourced from https://en.wikipedia.org/wiki/Motocross_des_Nations.

For this problem I came up with solution assuming that tidyr and ddplyr are not available and I have to do the tidy part myself.
   
I defined a function that pretty much performs the tidyr::gather function  
```{r}
my_gather=function(x,cols,keyname,valuename){
  all_keyvalue=NULL
  for (y in cols){c
    varname=colnames(x)[y]
    keyvalue=cbind(varname,x[,y,with=FALSE],x[,.I])
    colnames(keyvalue)=c(keyname,valuename,"row")#The extra row column is to keep track of the row to which this column needs to be attached to
    all_keyvalue=rbind(all_keyvalue,keyvalue)
  }
  all_keyvalue
  
}
```
   
Then, I performed the data cleansing and tidying as shown below  
```{r}

moto_wide=fread("https://raw.githubusercontent.com/mkds/MSDA/master/IS607/data/motocross.csv",stringsAsFactors = F)
head(moto_wide)
keyvalues=my_gather(moto_wide,2:17,"year","position")#Collapse columns to key value pair  
head(keyvalues)
#Extract Numbers from the strings
keyvalues[,`:=`(year=as.numeric(gsub("\n","",year)),position=as.numeric(gsub("(\\d+).*","\\1",position)))]
#Add row number as column so that we can join by row
moto_wide[,row:=.I]
setkey(moto_wide,row)
setkey(keyvalues,row)
#Join keyvalues with the required values from original data
moto=keyvalues[moto_wide[,.(Team,row)]][,.(Team,year,position)]
head(moto)

#Note that all above function could be performed easily using tidyr as given below
# moto=moto_wide %>% gather(year,position,2:17) %>% 
#      mutate(year=as.numeric(gsub("X..","",year)),position=as.numeric(gsub("(\\d+).*","\\1",position)))
#

#Now that we got the long data let's try to answer the question of  average finishes and average standing
kable(moto %>% 
  group_by(Team) %>% 
  summarize(finishes=sum(!is.na(position)),Avg_position=mean(position,na.rm=T)) %>%
  arrange(Avg_position))
