---
title: "Get Data Using Web API"
author: "Mohan Kandaraj"
date: "November 1, 2015"
output: html_document
---

This page demonstrates getting data from web using an Web API using an R program. New York Times provides web APIs that provide various content from nytimes.com. The Top Stories API is used in this program to fetch the top stories from NYT.

Three functions are developed to handle the Top Stories API.  
   
**Function: top_stories_uri**  
The function top_stories_uri constructs the URI for the top stories API given the section name and apikey
  
```{r}
#Function to construct URI
top_stories_uri=function(section,apikey,response_format=".json"){
  paste0("http://api.nytimes.com/svc/topstories/v1/",section,response_format,"?api-key=",apikey)
}
```  
  
**Function:process_top_stories***  
Function process_top_stories takes the json response from the top stories API and formats it into a dataframe   
   
```{r}
#Process json response
process_top_stories=function(response){
  library(jsonlite)
  #Convert json to nested dataframe
  response_dfs=fromJSON(response)
  #Get result count
  count = response_dfs$num_results
  #Initialize output data frame
  topstories=data.frame(matrix(numeric(0), nrow = count))
  results=response_dfs$results
  topstories$Title=results$title
  topstories$Abstract=results$abstract
  topstories$url=results$url
  topstories$By=results$byline
  topstories$Tags=results$org_facet
  #return output dataframe
  topstories
}
```
   
   
**Function: get_nyt_top_stories**  
This is the main function that act as the interface for the NYT top stories API. This function invokes the API for the given topic and returns formatted dataframe. This makes use of the other two functions mentioned above.
   
```{r.message=F,warning=F}

get_nyt_top_stories=function(section,apikey){
  library(RCurl)
  raw_response=""
  #Construct the URI for the request
  uri=top_stories_uri(section,apikey)
  raw_response=getURI(uri)
  #The GET failed, could be bad URI or bad connection
  if (raw_response==""){
    print("Error:Get Failed")
    return
  }
  #Able to get response, but response indicates it's error
  if (substr(raw_response,3,7)=="error"){
    print("Warning: Error response received")
    return
  }
  #Process the json response
  process_top_stories(raw_response)
  
}

```
   
**Let's try to get the top stories**  
  
```{r,message=F,warning=F}
#Read api key
apikey=as.character(read.table("nytapikey.txt",header = F,stringsAsFactors = F))
#Get top stories
topstories=get_nyt_top_stories("home",apikey)
library(knitr)
#display top stories
kable(topstories,caption="NYT Top Stories")
``` 
  
  
**Let's check the sports stories..**    
  
```{r,message=F,warning=F}
sportstories=get_nyt_top_stories("sports",apikey)
kable(sportstories)
```