---
title: "Recommender System for Music Recommendation"
output:
  html_document: 
    theme: cerulean
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,message = F,warning = F)
```

### Description 

The objective of this recommender system is to recommend music artists to users based on the previous music artists that the user often listened. The system uses collaborative filtering for recommendation. The number of time a user listened a particular artist is used as a rating (or proxy for rating). 


### Data 

The dataset used in this system contains the number of time a user listened music from a particular artist.  The dataset contains 92834 relations between 1892 users and 17632 artists. 

Data source: https://grouplens.org/datasets/hetrec-2011/ (Last.FM)

For this demonstration of the recommender system only popular artists and users who has enough relations are used.

#### Load Data

```{r}
#Read artist File
artists = read.table("artists.dat",sep="\t",fileEncoding = "UTF-8",stringsAsFactors = F,comment.char = "",quote="",header = T)
str(artists)
```

```{r}
user_artists = read.table("user_artists.dat",sep="\t",header = T)
str(user_artists)
```

### Data Transformation

#### Long to wide
The data provided is in long format. Let's convert to wider format such that each row represent the listened count of a user. 

```{r}
library(tidyr)
user_artists_wide = user_artists %>% spread(key=artistID,value=weight)
dim(user_artists_wide)
#Create character Id
artists$charid=paste0("I",artists$id)
```
```{r}
userids=user_artists_wide$userID
user_artists_wide$IuserID = NULL
rownames(user_artists_wide) = paste0("U",userids)
colnames(user_artists_wide) = paste0("I",colnames(user_artists_wide))
user_artists_wide[1:6,1:10]
```
### Filter, center and Scale data
In order to ensure we have enough data to work with and also to keep the computation cost down, the top 1000 artists and the users who has listened to at least 11 artists are kept in the dataset. 

In order to eliminate difference in scale of data between users the data is centered and scaled.

```{r}
#Select Top 1000 
visits_byitem=colSums(user_artists_wide[,-1],na.rm = T)

visits_1k = user_artists_wide[,order(visits_byitem,decreasing = T)[1:1000]]
num_visits=apply(visits_1k,1,function(x) return(sum(!is.na(x))))
visits_1k = visits_1k[num_visits>10,]
dim(visits_1k)
visits_1k=t(scale(t(visits_1k))[,])
```

### Methodlogy - User Based Collaborative Filtering

We shall use user based collaborative filtering method to come up with recommendation. The recommendation is first completely hand coded and then the results are compared with the recommenderlab package. As the methodology used in the hand coded method is same as the one used by recommenderlab the results of hand coded method is expected to be in-line with the results from recommederlab. 

**User based collaborative filtering**   

Recommendation for user is made based on similarity between users as described below  

* Compute similarity of the user with other user (using Pearson correlation)  
* Select top 50 similar user of the user for whom recommendation needs to be done  
* Compute items rating based on the recommendations done by the 50 similar users  
* Recommend items with high ratings    

#### Similarity Function
For similarity between users we shall use Pearson correlation. The function below provides similarity between user or products.


```{r}
#Function that computes similarity
similarity_user=function(all_data,user_data){
  #Transpose so that users are on columns
  all_data=t(all_data)
  #Use pearson correlation
  score = cor(all_data,user_data,use = "pairwise")
  return(score)
}
```

 

#### UBCF - Hand coded function

```{r}
#Function that predicts rating and list recommendation
ubcf_recommend=function(data,user_data,num_rec=20,num_sim=50){
  user_sim = similarity_user(data,user_data)
  #Replace NA with zero
  data_na0=data
  data_na0[is.na(data_na0)]=0
  top_sim_users=order(user_sim,decreasing = T)[1:num_sim]
  ratings=(user_sim[top_sim_users,1]%*%data_na0[top_sim_users,])
  prediction=NULL
  prediction$ratings=ratings
  #Set rating of already rated item to NA
  ratings[!is.na(user_data)]=NA
  prediction$recommendation=order(ratings,decreasing = T)[1:20]
  #To do remove already visited artists
  return(prediction)
}

```

### Split data for training and testing
In order to evaluate the recommendations the data is split into training and testing data. Training data is used for training and recommendations are computed for users in test data.

```{r}
library(recommenderlab)
visits_1k_rrm=as(as.matrix(visits_1k),"realRatingMatrix")
set.seed(100)
eval_sets <- evaluationScheme(data = visits_1k_rrm, method = "split", train = .8, given = 10, goodRating=3, k = 1)

```
Function to get artist name given the Id

```{r}
getArtistName=function(artistid){
  return(artists$name[artists$charid==artistid])
}
#Vectorize the function to enable get values for a vector of ids
getArtistName=Vectorize(getArtistName)
```  

### Get Recommendation using hand coded UBCF


```{r}
#Convert realRatingMatrix to regular Matrix
train=as.matrix(getRatingMatrix(getData(eval_sets,"train")))
test=as.matrix(getRatingMatrix(getData(eval_sets,"known")))
#realRating Matrix stores missing value as 0, convert them back to NA
train[train==0]=NA
test[test==0]=NA
#Check prediction for one user
pred1=ubcf_recommend(train,test[1,])
cat("\n**** Recommendation for user",rownames(test)[1],"\n")
recommend=getArtistName(colnames(train)[pred1$recommendation])
cat(recommend,sep="\n")

```

### Get Recommendation using recommenderlab

```{r}
model_ubcf=Recommender(data=getData(eval_sets,"train"),method="UBCF",
                          parameter=list(method="pearson",nn=50))
recommend_rl=predict(model_ubcf,getData(eval_sets,"known"),n=20)

cat("\n**** Recommendation for user",rownames(getData(eval_sets,"known"))[1],"\n")
recommend=getArtistName(recommend_rl@itemLabels[as.vector(recommend_rl@items[[1]])])
cat(recommend,sep="\n")

```


### Results Comparison  
The results obtained from the hand coded UBCF method is pretty much same as the result provided by recommenderlab package

```{r}
n = nrow(test)
recommendations=NULL
same_as_rl=logical(n)
for (i in 1:n){
  recommendations[[i]]=ubcf_recommend(train,test[i,])
  same_as_rl[i]=all(recommend_rl@items[[i]] %in% recommendations[[i]]$recommendation)
}
table(same_as_rl)

```

In 90% of the cases the recommendations made by hand coded function are exactly same as the recommendations made by recommenderlab package. In other instances the recommendations are very close to the recommendation made by the package and the difference could be due to implementation method/value approximation. 